---
title: "Explore and Summarize data project with R"
author: "Hindam"
date: "January 27, 2019"
output:
  html_document:
    include:
      in_header: github_corner.html
      after_body: footer.html
    css: style.css
    toc: TRUE
    toc_float: TRUE
    toc_depth: 3
    theme: flatly
    fig_height: 5
    fig_width: 9
    highlight: tango
      
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE ,message=FALSE, warning=FALSE)
```

```{r}
# call the required packages 
library(ggplot2)
library(psych)
library(dplyr)
library(GGally)
library(gridExtra)
```


This report explores a dataset containing quality and attributes for approximately 1600 red wines.
***

```{r}
# read data file
df <- read.csv("wineQualityReds.csv")
dim(df)
```

***

- So, Our data consists of 1599 observations and 13 variables
```{r}
# invistigate variables
str(df)
```

***

```{r}
# data descriptive statistics summary
summary(df)
```

***

# Univariate Plots Section

- First as the quality is our main concern, let's see the distribution of quality

```{r}
# Histogram - Quality
g1 <- ggplot(aes(x= as.factor(quality)), data= df)+
  geom_histogram(stat = 'count')
g1
```

- From the above plot we can notice that the distribution of quality is **normal** and also **bimodal**.
- Most of the red wine quality at 5 & 6 grades no quality below 3 and there is no quality higher than 8.
- Now, Let's plot simple historams of all other variables to see their distribution

***

```{r}
# Histogram - fixed.acidity
ggplot(aes(x= fixed.acidity), data= df)+
  geom_histogram()
```
```{r}
summary(df$fixed.acidity)

```

- Here the distribution of fixed acidity is to some extent **normally distributed** 
- The majority of fixed acidity between 6-9.
- PPresence of outlier as high as 15.9 

***

```{r}
# Histogram - volatile.acidity
ggplot(aes(x= volatile.acidity), data= df)+
  geom_histogram()
summary(df$volatile.acidity)

```

- The distribution of volatile acidity is considered **normal** with _two beaks_
- The mean and median values are close
- The majority between _0.2_ and _0.8_ and some outliers are present

***

```{r}
# Histogram - citric.acid
ggplot(aes(x= citric.acid), data= df)+
  geom_histogram()
```
```{r}
summary(df$citric.acid)
```

- There are some red wines with zero citric acid and there is still some outliers
- The distribution here is **positively skewed** with _two peaks_
- The mode of is _0_, which is consistent with the data info " found in small quantities, citric acid can add _freshness_ and _flavor_ to wines."

***

```{r}
# Histogram - residual.sugar
ggplot(aes(x= residual.sugar), data= df)+
  geom_histogram()
```
```{r}
summary(df$residual.sugar)
```

- The distribution is **Right skewed** 
- The majority below _2.6_ with some outliers which is considered **'Sweet*'** in this case.

***

```{r}
# Histogram - chlorides
ggplot(aes(x= chlorides), data= df)+
  geom_histogram()

```

```{r}
# Summary - chlorides
summary(df$chlorides)
```

- Some outliers above **0.6** while the majority less than **0.1**
- All red wines in this dataset has a proportion of chlorides.

***

```{r}
# Histogram - Free SO2
ggplot(aes(x= free.sulfur.dioxide), data= df)+
  geom_histogram()
# Smmary - Free SO2
summary(df$free.sulfur.dioxide)

```

- The distribution here is **Right skewed** and with the most frequent value _Mode_ around 5.
- Presence of vertical peaks indicates common concentration usd in different red wines.
- 75% of the datapoits have free SO2 concentration below _21_ and there outliers as high as _72_ 
   
***

```{r}
# Histogram - Total SO2
ggplot(aes(x= total.sulfur.dioxide), data= df)+
  geom_histogram()
# summary - Total SO2
summary(df$total.sulfur.dioxide)
```

- The distribution again here is **Right Skewed**
- Too big values for the outliers while nearly _75%_ of the values below _62_

***

```{r}
# Histogram - density
ggplot(aes(x= density), data= df)+
  geom_histogram()
```

- Density distribution takes a **bell curve** style with the majority between 0.995 and 1

***

```{r}
# Histogram - pH
ggplot(aes(x= pH), data= df)+
  geom_histogram()
```

- Most PH values are between 3 & 3.5 with **Normal Distribution**

***

```{r}
# Histogram - Sulphates
ggplot(aes(x= sulphates), data= df)+
  geom_histogram()
# Summary - Sulphates
summary(df$sulphates)
```
- The distribution of sulphates in our dataset is **Right skewed**
- The median and mean values are very close
- The majority of red wines in this dataset has been in the range of _0.5 to 1_
- Again presence of some outliers

***

```{r}
# Histogram - Alcohol
ggplot(aes(x= alcohol), data= df)+
  geom_histogram()
# Summary - Alcohol
summary(df$alcohol)
```

- Alcohol distibution is **Right skewed** with very close values for _mean_ and _median_
- The avearge concentration of Alcohol here is around 10%
- Some outlier values above 14, _max = 14.90_ while the minimum alcohol concentration at _8.4%_
- The most frequent value a little above _9_

***  


```{r echo=FALSE}
# make a new variable for rating based on the quality of red wines
df$rating = cut(df$quality,breaks = 3 , labels = c("Bad", "Average","Good"))
# Summary - rating
summary(df$rating)
```

```{r}
# distribution - rating
g2 <- ggplot(aes(x= rating), data = df)+
  geom_bar()
g2
```

- From the above we conclude the majority of red wines rating is **average**

***

# Univariate Analysis


#### What is the structure of your dataset?

- There are 1599 wines in the dataset with 13 features.
X referes for indexing.
- The variable of our intrest is **quality** and we generate **rating**
- 11 _chemical_ variables : 
(fixed.acidity,volatile.acidity,citric.acid,residual.sugar,chlorides, free.sulfur.dioxide,total.sulfur.dioxide,density,pH,sulphates,alcohol)

#### What is/are the main feature(s) of interest in your dataset?

- The main feature here is **quality**. 
- I would dive in the relationships between quality and other variables.

#### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

- From my point of view I think the following variables could share for wine quality:
 1. The Alcohol 
 2. Volatile acidity as its high values leads to an unpleasant, vinegar taste
 3. Citric acid as its the cause of freshness and flavors to wine
 4. residual sugar as it affects the taste.
 
#### Did you create any new variables from existing variables in the dataset?

- Yes, I create a vriable **rating**.which is ordered factor with levels of **bad, average and good**

#### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

- From the above plots I noticed some normal distribution and some Right skewed:
  1. **Normal distribution: **  fixed.acidity, volatile.acidity, density and PH
  2. **Right Skewed :** citric.acid,cholrides,total.sulfur.dioxide, and sulphates.
- There is no tidy or adjust work as the data comes tidy, only I added the variable rating, to   figure out the quality effect.

***


# Bivariate Plots Section

- First let's plot the correlation between our intested features.

```{r }
# correlation matrix for selected 5 variables
mydata <- df%>%
   select(alcohol,volatile.acidity,citric.acid,residual.sugar,quality)

# plot 
pairs.panels(mydata, scale=TRUE)



```

- Here we can notice that both alcohol and volatile.acidity are moderately correlated with quality which has almost no relationship with residual.suagar
- Almost strong negative correlationship between citric.acid and volatile.acidity

***

- Now let's plot all the features as we could find something new

```{r wider_plot,fig.height=7,fig.width=9}
# correlation matrix for all variables
ggcorr(df[ , -c(1,14)], nbreaks=8, palette='RdGy', label=TRUE, label_size=5, label_color='white', hjust = 0.8, # Horizontally Adjust text
       layout.exp = 2) # Expand the layout
  
```

- From the above plot we can notice some strong realtionships between features like pH and acidity but my aim here is to focus on **qaulity**.

#### Correlation:

1. alcohol (moderate positive correlation with quality )
2. volatile acidity (weak negative correlation with quality)
3. sulphates (moderate positive correlation with quality )

- So, We can add sulphates into consideration.
-  I want to dive in scatter plots for _quality_ and some other variables like _alcohol_, _sulphates_, and _volatile acidity_.

***

```{r}
# convert quality to factor variable
df$quality <- factor(df$quality, ordered = T)

# Box plot - Quality & Alcohol
g3<-ggplot(data = df, aes(x = quality, y = alcohol,fill = rating)) +
  geom_boxplot() +
  ggtitle('Alcohol in Different Red Wine Qualities') +
  xlab('Quality') +
  ylab('Alcohol (% by volume)')
g3

# group by rating to calculate Alcohol median
rating_alc <- df %>%
  group_by(rating) %>%
  summarise(alcohol_median = median(alcohol),
            n=n())
rating_alc
```

- From the above plot we can notice that **good quality** red wines has the **largest median** Alcohol compared to the other qualities, it's obvious.

***

```{r}
# Box plot - Quality & Sulphates
ggplot(data = df, aes(x = quality, y = sulphates,fill = rating)) +
  geom_boxplot() +
  ggtitle('Sulphates in Different Red Wine Qualities') +
  xlab('Quality') +
  ylab(expression(Sulphates~(potassium~sulphates~g/dm^{3})))
```


```{r}
# group by rating to calculate sulphates median
rating_sulph <- df %>%
  group_by(rating) %>%
  summarise(sulphates_median = median(sulphates),
            n=n())
rating_sulph
```
- Here too we can notice that **good quality** red wines has the largest **median** sulphates compared to the other qualities.

***


```{r}
# Box plot - Quality & volatile.acidity
ggplot(data = df, aes(x = quality, y= volatile.acidity, fill= rating)) +
  geom_boxplot()+
  ggtitle('Volatile acidity in Different Red Wine Qualities') +
  xlab('Quality') +
  ylab( expression(Volatile~acidity~(acetic~acid~g/dm^{3})))

# group by rating to calcultae volatile.acidity median
rating_vol <- df %>%
  group_by(rating) %>%
  summarise(volatile_acidity_median = median(volatile.acidity),
            n=n())
rating_vol
```

- Here too as we expect  that **good quality** red wines has the lowest **median** volatile.acidity compared to the other qualities.

***

Now let's discover the other variables.

```{r}
# Box plot - Quality & fixed.acidity
ggplot(data = df, aes(x = quality, y= fixed.acidity, fill= rating)) +
  geom_boxplot()+
  ggtitle('Fixed acidity in Different Red Wine Qualities') +
  xlab('Quality') +
  ylab(expression(Fixed~acidity~(tartaric~acid~g/dm^{3})))+
  geom_point(stat='summary', fun.y= mean, shape= 4)

# group by rating to calcultae fixed.acidity median
rating_fix <- df %>%
  group_by(rating) %>%
  summarise(fixed_acidity_median = median(fixed.acidity),
            n=n())
rating_fix
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

- So, from the above it seems that **higher quality** wines has the **largest median fixed acidity**, while the median for both average and bad are so close, with a high range in the average quality,it was not obvious.

</div>
***


```{r}
# Box plot - Quality & citric.acid
 ggplot(data = df, aes(x = quality, y= citric.acid, fill= rating)) +
  geom_boxplot()+
  ggtitle('Citric acid in Different Red Wine Qualities') +
  xlab('Quality') +
  ylab(expression(Citric~acid~(g/dm^{3})))

```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">


- Another noticable trend here although it was weak correlation between citric acid and quality, but from the above plot we can notice that **higher quality** red wines has the **largest median citric acid** while the lowest median citric acid in lowest quality.

</div>
***


```{r}
# Box plot - Quality & residual.sugar
ggplot(data = df, aes(x = quality, y= residual.sugar, fill= rating)) +
  geom_boxplot()+
  ggtitle('Residual Sugar in Different Red Wine Qualities') +
  xlab('Quality') +
  ylab(expression(Residual~Sugar~(g/dm^{3})))
```

- median residual sugrar is almost the same, with presence of outliers , Although I assumed the importance of residual sugar effect on quality but from the above plot I get back. 

***

```{r}
# Box plot - Quality & chlorides
ggplot(data = df, aes(x = quality, y= chlorides, fill= rating)) +
  geom_boxplot()+
  ggtitle('Chlorides in Different Red Wine Qualities') +
  xlab('Quality') +
  ylab(expression(Chlorides~(sodium~chloride~g/dm^{3})))
```

- Almost the same median of chlorides for all qualities, except the smalest range for the highest quality, also outlier present here.

```{r}
# group by rating to calculate Chlorides median
rating_chlor <- df %>%
  group_by(rating) %>%
  summarise(chlorides_median = median(chlorides),
            n=n())
rating_chlor
```

- Here higher quality wines have the lowest chlorides median with minor difference.

***

```{r}
# Box plot - Quality & Free SO2
ggplot(data = df, aes(x = quality, y= free.sulfur.dioxide, fill= rating)) +
  geom_boxplot() +
  ggtitle('Free SO2 in Different Red Wine Qualities') +
  xlab('Quality') +
  ylab(expression(Free~SO2~(mg/dm^{3})))
```

```{r}
# Box plot - Quality & Total SO2
ggplot(data = df, aes(x = quality, y= total.sulfur.dioxide, fill= rating)) +
  geom_boxplot() +
  ggtitle('Total SO2 in Different Red Wine Qualities') +
  xlab('Quality') +
  ylab(expression(Total~SO2~(mg/dm^{3})))

```

- From the above two plots for Free SO2 and total SO2, it seems that the largest median is present in the avearage rating group.

***

```{r}
# Box plot - Quality & Density
ggplot(data = df, aes(x = quality, y= density, fill= rating)) +
  geom_boxplot() +
  ggtitle('Density in Different Red Wine Qualities') +
  xlab('Quality') +
  ylab(expression(Density~(g/cm^{3})))

```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

- Another notice here as it was weak negative correlation between density and quality, but from the above plot we can be decieved by a misleading trend that **higher quality** red wines has the **lowest median density** and the highest density in the lowest quality red wines.that is beacuse the y-axis not start from 0, but the actually here all red wines have close density in range 0.995-1 which is also close to water

</div>


***

```{r}
# Box plot - Quality & pH
ggplot(data = df, aes(x = quality, y= pH,fill= rating)) +
  geom_boxplot() +
  ggtitle('pH in Different Red Wine Qualities') +
  xlab('Quality') +
  ylab('pH')

```

- From the above we can notice a trend between pH and quality,as in the highest quality of red wines pH medain values are the lowest.

***
Now let's figure out the realtion between different variables

```{r}
# Scatter plot - different variales 
p1 <- ggplot(aes(x = pH, y = fixed.acidity),data=df)+
  geom_point(alpha= 0.1, position = position_jitter(h=0))+
  geom_smooth(method ='lm', color='purple')

p2 <- ggplot(aes(x = pH, y = citric.acid), data=df)+
  geom_point(alpha= 0.1, position = position_jitter(h=0))+
  geom_smooth(method ='lm', color='purple')

p3 <- ggplot(aes(x = pH, y = volatile.acidity), data=df)+
  geom_point(alpha= 0.1, position = position_jitter(h=0))+
  geom_smooth(method ='lm', color='orange')

p4 <- ggplot(aes(x=citric.acid, y= volatile.acidity),data=df)+
  geom_point(alpha= 0.1, position = position_jitter(h=0))+
  geom_smooth(method ='lm', color='purple')

p5 <- ggplot(aes(x=citric.acid, y= fixed.acidity),data=df)+
  geom_point(alpha= 0.1, position = position_jitter(h=0))+
  geom_smooth(method ='lm', color='orange')

p6 <- ggplot(aes(x= density, y = fixed.acidity),data=df)+
  geom_point(alpha= 0.1, position = position_jitter(h=0))+
  geom_smooth(method ='lm', color='orange')


grid.arrange(p1 , p2, p3, p4, p5, p6, ncol=3)

```

- It worth to figure how these variables correaltes to each other:
1. **Postiviely**: 
                   -pH with volatile.acidity
                   -fixed acidity with both density and citric.acid.
2. **Negatively**: 
                  -pH with both citric.acid and fixed.acidity
                  -citric.acid with volatile acidity.

***

# Bivariate Analysis

#### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

- Quality correlates strongly with alcohol, volatile acidity and sulphates.
- Red wines with high _alcohol_, _sulphates_, _fixed acidity_ and _citric acid_ tend to have **higher quality**
- Red wines with low _volatile acidity_, _density_ and _pH_ tend to have **higher quality**
- **Average quality red wines** have the highest _Free S02 & Total SO2._ This seems really unusual since I would expect *bad quality*  to have a higher median _Free S02 & Total SO2._ compared to the other groups which may affect their taste in the test. 

#### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

1. **Positive Correlations** :
                   - Alcohol with quality
                   - pH with volatile.acidity
                   - fixed acidity with both density and citric.acid.
2. **Negative Correlations**: 
                  - pH with both citric.acid and fixed.acidity
                  - citric.acid with volatile acidity.

#### What was the strongest relationship you found?
1. **Strongest Positive Correlations** : Alcohol with quality.
2. **Strongest Negative Correlations**: pH with with and fixed.acidity
                  

***

# Multivariate Plots Section

From the previous bivariate analysis we conclude that Alcohol has the strongest correlation
wih quality.

```{r}
# Distribution - Quality & Alcohol
ggplot(aes(x= quality , y =  alcohol),
       data = df)+
  geom_bar(aes(fill = rating), stat = 'summary', fun.y= median)+
  ggtitle('Median of Alcohol in Different Red Wine Qualities')+
  ylab('Median Alcohol (% by volume)')

# group by rating & quality to calculate Alcohol mean & median
df %>%
 
  group_by(rating, quality) %>%
  summarise(mean_alc = mean(alcohol),
            median_alc= median(alcohol),
            n= n()) %>%
  ungroup() %>%
  arrange(rating)
```

The median for alcohol concentration is increasing across quality levels.

***

Now, Let's divide in more about alcohol with other variables starting with fixed acidity.

```{r}
# Scatter plot - Alcohol & fixed.acidity across different rating
# use color for rating
ggplot(aes(x= alcohol, y= fixed.acidity,color = rating),
       # exclude average rating
       data= subset(df,rating != 'average'))+
  geom_smooth()+
  geom_point(alpha = 0.8)+
  ggtitle('Fixed acidity VS Alcohol in Red wines')+
  scale_color_brewer(palette = "RdYlBu", name="Rating", direction=-1)+
  xlab('Alcohol (% by volume)')+
  ylab(expression(Fixed~acidity~(tartaric~acid~g/dm^{3})))


```

- The higher quality of red wines tend to have **higher fixed acidity**.


```{r}
# Scatter plot - Alcohol & volatile.acidity across different rating
# use color for rating 
# exclude average rating
ggplot(aes(x= alcohol, y= volatile.acidity,color = rating),
       data= subset(df,rating != 'average'))+
  geom_smooth()+
  geom_point(alpha = 0.8)+
  ggtitle('Volatile acidity VS Alcohol in Red wines') +
  scale_color_brewer(palette = "RdYlBu", name="Rating", direction=-1)+
  xlab('Alcohol (% by volume)')+
  ylab(expression(Volatile~acidity~(acetic~acid~g/dm^{3})))

```

- On the other hand tend to have a lower volatile acidity , so let's figure how they act in the ratio of fixed:volatile acidity

```{r}
# Scatter plot - Alcohol & fixed.acidity/volatile.acidity ratio 
# with facet all rating group 
# use color for quality 
ggplot(aes(y= fixed.acidity/volatile.acidity, x= alcohol),
       data= df)+
  geom_point(aes(color= quality),
            stat = 'summary',
            fun.y= median)+
  facet_grid(~rating)+
  ggtitle('Fixed acidity/ volatile acidity VS Alcohol in Red wines',
          subtitle = 'Facet Wrap by rating')+
  scale_color_brewer(palette = "RdYlBu", name="Quality", direction=-1)+
  xlab('Alcohol (% by volume)')+
  ylab(expression(frac(Fixed~Acidity~(~tartaric~acid~g/dm^{3}),
                       Volatile~Acidity~(acetic~acid~g/dm^{3}))))

```

- Let's focus more on difference between bad and good qualities.

```{r}
# Scatter plot - Alcohol & fixed.acidity/volatile.acidity ratio
# for bad & good rating setting horizontal limit
# use color for rating 

g4 <-ggplot(aes(x= alcohol, y=  fixed.acidity/volatile.acidity ,color = rating ),
       data= df)+
  geom_smooth()+
  geom_jitter(alpha = 0.8)+
  geom_hline(yintercept = 20, linetype= 5)+
  ggtitle('Fixed acidity/ volatile acidity VS Alcohol in Red wines') +
  scale_color_brewer(palette = "RdYlBu", name="Quality", direction=-1)+
  xlab('Alcohol (% by volume)')+
  ylab(expression(frac(Fixed~Acidity~(tartaric~acid~g/dm^{3}),
                       Volatile~Acidity~(acetic~acid~g/dm^{3}))))

g4
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

- That is intersting for me, the higher qulaity red wines tend to have higher fixed acidity to volatile acidity ratio above _20_, also the majority of average & bad red wines at different concentration of alcohol tends to have fixed:volatile ratio at below that line.

</div>

***

```{r}
# Scatter plot - Alcohol & sulphates across different rating
# use facet wrap for all rating
# use color for quality 
ggplot(aes(x= alcohol, y= sulphates,color = quality ),
       data= df)+
  geom_point(alpha=0.8)+
  facet_grid(~rating)+
  geom_smooth()+
  ggtitle('Sulphates VS Alcohol in Red wines',
          subtitle = 'Facet Wrap by rating')+
  scale_color_brewer(palette = "RdYlBu", name="Quality", direction=-1)+
  xlab('Alcohol (% by volume)')+
  ylab(expression(Sulphates~(potassium~sulphates~g/dm^{3})))
  
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

- Although high quality red wines tend to have higher sulphates conc., but the effect of sulphates here is not quiet clear to determine its effect on quality specially for the higher variability in the avearge group.

</div>

***


```{r}
# Scatter plot - Alcohol & citric.acid across different rating
# use color for rating 
# exclude average rating
ggplot(aes(x= alcohol , y =  citric.acid, color= rating),
       data = df)+
  geom_point(alpha=0.8)+
  geom_smooth()+
  ggtitle('Citric acid VS Alcohol in Red wines')+
  scale_color_brewer(palette = "RdYlBu", name="Rating", direction=-1)+
  xlab('Alcohol (% by volume)')+
  ylab(expression(Citric~acid~(g/dm^{3})))

```

- Although citric acid is resposible for freshness at low concentration, but as per the above plot I noticed that most of higher quality red wines have a higher concentration of citric acid and the majority at 0 for bad quality.

***

```{r}
# Scatter plot - Alcohol & pH across different rating
# use color for rating 
ggplot(aes(x= alcohol , y = pH, color = rating),
       data = df)+
  geom_point(alpha=0.8)+
  geom_smooth()+
  ggtitle('pH VS Alcohol in Red wines')+
  scale_color_brewer(palette = "RdYlBu", name="Rating", direction=-1)+
  xlab('Alcohol (% by volume)')
```

- Although the correlation between quality and pH was _-0.1_ , but here high quality wines has tendency for lower pH specially less than _3.2_ which sounds logic as higher quality alcohol tend to have more citric acid and fixed acidity which for sure leads to lower pH values.

***

```{r}
# Scatter plot - Alcohol & total acidity
# total acidity = fixed+volatile+citric 
# use color for rating 
# exclude average rating
ggplot(aes(x= alcohol , y = (citric.acid+fixed.acidity + volatile.acidity), color= rating),
       data = subset(df,rating !='average'))+
  geom_point(alpha=0.8)+
  geom_smooth()+
  ggtitle('Total acidity VS Alcohol in Red wines')+
  ylab('Total acidity')+
  scale_color_brewer(palette = "RdYlBu", name="Rating", direction=-1)+
  xlab('Alcohol (% by volume)')+
  ylab(expression(Total~acidity~(g/dm^{3})))
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

- Intersting point here as although higher quality red wines tend to have less volatile acidity but the overall comined acidity in them is higher which reflects the pH values to be smaller in compasion to bad quality ones.

</div>

```{r}
# Scatter plot - Alcohol & Free SO2:Bound SO2 ratio
# Bound SO2 = total SO2- Free SO2
# use facet wrap for all rating
# use color for quality 

ggplot(aes(y= (free.sulfur.dioxide/(total.sulfur.dioxide - free.sulfur.dioxide)), 
           x=  alcohol ,color = rating ),data= df)+
  geom_point(alpha=0.8)+
  geom_smooth()+
  ggtitle('Free SO2 / Bound SO2 VS Alcohol in Red wines',
          subtitle = 'Facet Wrap by rating') +
  ylab('ratio')+
  facet_wrap(.~rating)+
  scale_color_brewer(palette = "RdYlBu", name="Rating", direction=-1)+
  xlab('Alcohol (% by volume)')+
  ylab(expression(frac(Free~SO2~(mg/dm^{3}),Bound~SO2~(mg/dm^{3}))))


```

- As a try to calculate the ratio of free SO2 to bound SO2, it seems that all the quality levels behave in the same trend with majority having ratio less than 1 except for the variance in the average group.

***

# Multivariate Analysis

#### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?
High quality red wines have the largest median & mean for alcohol concentration.

In comparing to alcohol, I found that high quality alcohols tends to have higher fixed acidity and citric acid, but lower volatile acidity and pH.

No effect for sulphates neither nor ration of Free SO2: Bound SO2.

#### Were there any interesting or surprising interactions between features?
By measuring Fixed acidity to volatile acidity ration, I noticed that the higher qulaity red wines tend to have higher ratio above 20.

Despite the positive correlation between quality and sulphates but I conclude that sulphates acts in different manner in compaing to alcohol across all rating levels

Although the correlation between quality and pH was too weak and negative , but high quality wines has tendency for lower pH specially less than 3.2.

Citric acid is higher in high quality red wines than in bad quality.

***

# Final Plots and Summary


### Plot One
```{r}
# Plot 1: - Histogram - Quality/ rating
plot1 <- g1+
  xlab('quality')+
  ggtitle('Red Wines Distribution by Quality')

plot2 <- g2+
  ggtitle('Red Wines Distribution by rating')

grid.arrange(plot1,plot2,ncol=2)

```

### Description One
The histogram shows clearly that most of the red wines quality on sclaes 5 & 6
and so they have the average rating.


***
<div class = "row">
  
<div class = "col-md-8">
<h3>Plot Two</h3>
```{r}
# Plot 2: Box plot - Alcohol & Quality
g3

```
</div>

<div class = "col-md-4">
<h3>Description Two</h3>
<br><br>The boxplots here show good quality red wines has the largest median Alcohol compared to the other qualities, it's obvious..
</div>

</div>

***

<div class = "row">

<div class = "col-md-8">
<h3>Plot Three</h3>
```{r}
# Plot 3: Scatter plot - Alcohol & fixed.acidity/volatile.acidity ratio
# for bad & good rating setting horizontal limit
# use color for rating 
# exclude average rating
g4

```
</div>

<div class = "col-md-4">
<h3>Description Three</h3>
<br><br>The higher qulaity red wines tend to have higher fixed acidity to volatile acidity ratio above 20, also the majority of average & bad red wines at different concentration of alcohol tends to have fixed:volatile ratio at below that line.
</div>

</div>

***
# Reflection
The red wines data set contains information on almost 1,599 red wines across 11 variables from around 2009. I started by understanding the individual variables in the data set, and then I explored interesting questions and leads as I continued to make observations on plots. Eventually, I explored the quality of diamonds across many variables.

There was a clear trend between the alcohol concentration or volatile acidity of a red wine and its quality. I was surprised that citric acid did not have a strong positive correlation with quality. I struggled understanding the variance in quality as the level of  Free S02 & Total SO2 get increased, but this became more clear when I realized that most of the data contained average rating of quality rank 5 &6.

Some limitations of this data include:

- The source of the data. Given that the red wines date to 2009.
- The imbalance nature of the data most of the quality related to the average group.
- It's related to only one brand , quality could be differed in presence of other brand
- I have not expeeinced before with wine, so I guess in this topic domain experience is usefull, and this could suggest the idea of using addictive wine user for the test.
- As per my google search people generally have individual preference for wines, even that     each occassion has its own wine atmosphere and kind, with meal or without, party or gift.
- Some other variables could be added like tannins content.

***

# Resources

<ul>
<li><a href="'https://holtzy.github.io/Pimp-my-rmd/'">Pimp my RMD</a></li>
<li>The <a href="'https://rmarkdown.rstudio.com/'">R markdown website</a></li>
<li><a href="'http://jamesmarquezportfolio.com/correlation_matrices_in_r.html'">Correlation plots</a></li>
<li><a href="'http://tholman.com/github-corners/'">Github corner</a></li>
</ul>

```{r}

```

